---
title: "data_with_random_intercept"
output:
  html_document: default
  pdf_document: default
date: "2024-02-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library
```{r library}
library(MASS) # For mvrnorm to generate multivariate normal distributions
library(GeneralisedCovarianceMeasure) # Generalised Covariance Measure (GCM)
library(CondIndTests) # kernel-based CI test (KCI)
library(scales)
library(Matrix)
library(MASS)
library(lme4)
library(mgcv)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gamm4)
library(lmtest)
```

# Linear model with random intercept

X_it= 1.5*Z_t +alpha_i+e_it
Y_it= 1.1*Z_t +beta_i+e_it

```{r dataset of linear case}
set.seed(123)

b <- 0

n_subjects <- 20
n_measurements <- 200

cor_matrix <- matrix(0.5, nrow = n_measurements, ncol = n_measurements)
diag(cor_matrix) <- 1

data <- data.frame(subject = rep(1:n_subjects, each = n_measurements),
                   measurement = rep(1:n_measurements, times = n_subjects),
                   Z = NA)

# Generate Z values
for (i in 1:n_subjects) {
  data$Z[data$subject == i] <- mvrnorm(n = 1, mu = rep(0, n_measurements), Sigma = cor_matrix)
}

# Initialize X and Y with NA values
data$X <- NA
data$Y <- NA

# Function to calculate X based on current and previous Z
calculate_X_Y <- function(Z_current, Z_previous, is_X = TRUE) {
  # Handle the case for the first measurement
  Z_previous <- ifelse(is.na(Z_previous), Z_current, Z_previous)
  
  if(is_X) {
      return(Z_current)
    # return(5 * sin((Z_current + Z_previous)^2 + 10))
  } else {
    return(Z_current)
    #return((Z_current + Z_previous)^2)
  }
}


alpha <- rnorm(n_subjects,sd=5)
alpha <- rep(alpha, each = n_measurements)

beta <- rnorm(n_subjects,sd=5)
beta <- rep(beta, each = n_measurements)

# Iterate through data to calculate X and Y based on Z_t and Z_t-1
for(i in 1:nrow(data)) {
  Z_current <- data$Z[i]
  Z_previous <- ifelse(data$measurement[i] == 1, NA, data$Z[i-1])
  data$X[i] <- calculate_X_Y(Z_current,0, TRUE) + alpha[i] + rnorm(1, mean = 0, sd = 1)
  data$Y[i] <- calculate_X_Y(Z_current,0, FALSE) + beta[i] + b*data$X[i]  +rnorm(1, mean = 0, sd = 1)
}
```

# Check data structure
```{r 1Plot of X,Y on Z}
library(ggplot2)


ggplot(data, aes(x = Z, y = X)) +
  geom_point(aes(color = Z), alpha = 0.6) +  
  scale_color_gradient(low = "blue", high = "red") + 
  theme_minimal() +  
  labs(title = "Relationship between Z and X",
       x = "Z",
       y = "X") +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Z, y = Y)) +
  geom_point(aes(color = Z), alpha = 0.6) +  
  scale_color_gradient(low = "green", high = "yellow") + 
  theme_minimal() +  
  labs(title = "Relationship between Z and Y",
       x = "Z",
       y = "Y") +
  theme(plot.title = element_text(hjust = 0.5))  
```
```{r Average X and Y vs. Average Z}
# Use tidyverse to calculate averages and plot
data %>%
  group_by(subject) %>%
  summarise(average_X = mean(X, na.rm = TRUE), average_Z = mean(Z, na.rm = TRUE), average_Y = mean(Y, na.rm = TRUE)) %>%
  ggplot() +
  geom_point(aes(x = average_Z, y = average_X), color = "blue") +
  geom_point(aes(x = average_Z, y = average_Y), color = "red") +
  labs(x = "Average Z", y = "Average", title = "Average X and Y vs. Average Z for each subject") +
  theme_minimal() +
  scale_y_continuous(name = "Average X", sec.axis = sec_axis(~., name = "Average Y"))

```
```{r Relationship between Z and Y for Selected Subjects}
library(ggplot2)

set.seed(123)
selected_subjects <- sample(unique(data$subject), 10)
selected_data <- data[data$subject %in% selected_subjects,]


ggplot(selected_data, aes(x = Z, y = X, group = subject, color = as.factor(subject))) +
  geom_line() +  
  geom_point() +  
  theme_minimal() +
  labs(title = "Relationship between Z and X for Selected Subjects",
       x = "Z",
       y = "X",
       color = "Subject") +
  scale_color_viridis_d(name = "Subject ID") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 10))

ggplot(selected_data, aes(x = Z, y = Y, group = subject, color = as.factor(subject))) +
  geom_line() +  
  geom_point() +  
  theme_minimal() +
  labs(title = "Relationship between Z and Y for Selected Subjects",
       x = "Z",
       y = "Y",
       color = "Subject") +
  scale_color_viridis_d(name = "Subject ID") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 10))

```

# Generalized additive regression

```{r Generalized additive regression}
gam_model_x <- gam(X ~ s(Z), data = data, method = "GCV.Cp")


gam_model_y <- gam(Y ~ s(Z), data = data, method = "GCV.Cp")


ggplot(data, aes(x = Z, y = X)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "gam", formula = y ~ s(x), method.args = list(family = "gaussian")) +
  ggtitle("Relationship between X and Z") +
  theme_minimal()


ggplot(data, aes(x = Z, y = Y)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "gam", formula = y ~ s(x), method.args = list(family = "gaussian")) +
  ggtitle("Relationship between Y and Z") +
  theme_minimal()
```

```{r Durbin Watson Test}
dwtest(gam_model_x)
dwtest(gam_model_y)
```


```{r Residuals by Subject 1}
data <- data %>% arrange(subject, Z)
data$residuals_x <- residuals(gam_model_x)
data$residuals_y <- residuals(gam_model_y)

ggplot(data, aes(x = Z, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_line() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X vs. Z by Subject with Lines") +
  theme_minimal() +
  labs(colour = "Subject")


ggplot(data, aes(x = Z, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_line() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y vs. Z by Subject with Lines") +
  theme_minimal() +
  labs(colour = "Subject")
```

```{r Residual plot 1}
data$index <- 1:nrow(data)
ggplot(data, aes(x = index, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject") 


ggplot(data, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject") 
```

```{r Correlation of residual}
res_x <- comp.resids(data$X, data$Z, regr.pars = list(), regr.method = "gam")
res_y <- comp.resids(data$Y, data$Z, regr.pars = list(), regr.method = "gam")
acf(res_x)
acf(res_y)
```

# GCM based on GAM

```{r GCM with GAM 1}
pvals_gcm <- gcm.test(data$X, data$Y, data$Z, regr.method = "gam", plot.residuals = TRUE, regr.pars = list())$p.value

# Interpretation
cat("Based on the performed GCM test:\n")
cat("- With a p-value of", pvals_gcm, "and a significance level of 0.05,\n")
if (pvals_gcm < 0.05) {
  cat("- We reject the null hypothesis.\n")
} else {
  cat("- We fail to reject the null hypothesis.\n")
}
```
# Linear mixed effects model
```{r}
# Model with Random Intercepts
# accounting for variations across groups (e.g., individuals) by allowing intercepts to vary.

model_X_lme_1 <- lmer(X ~ Z + (1|subject), data = data)
model_Y_lme_1 <- lmer(Y ~ Z + (1|subject), data = data)


residuals_X_lme_1 <- resid(model_X_lme_1)
residuals_Y_lme_1 <- resid(model_Y_lme_1)
```

```{r Residual plot & ACF}
plot(residuals_X_lme_1)
abline(h = 0, col = "red")
acf(residuals_X_lme_1)
plot(residuals_Y_lme_1)
abline(h = 0, col = "red")
acf(residuals_Y_lme_1)

```
```{r LME Model Predictions for Each Subject}
library(lme4)
library(ggplot2)
library(dplyr)



predict_data <- expand.grid(
  Z = seq(min(data$Z), max(data$Z), length.out = 100),  # 假设Z的范围
  subject = unique(data$subject)
)


predict_data$predicted_X <- predict(model_X_lme_1, newdata = predict_data, re.form = NULL)


ggplot() +
  geom_point(data = data, aes(x = Z, y = X, colour = factor(subject)), alpha = 0.5) + 
  geom_line(data = predict_data, aes(x = Z, y = predicted_X, colour = factor(subject)), size = 1) + 
  labs(title = "LME Model Predictions for Each Subject", x = "Z", y = "X") +
  theme_minimal() +
  theme(legend.title = element_blank())  


```

```{r Residuals by Subject}
data <- data %>% arrange(subject, Z)
data$residuals_x <- residuals(model_X_lme_1)
data$residuals_y <- residuals(model_Y_lme_1)

ggplot(data, aes(x = Z, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X vs. Z by Subject with Lines") +
  theme_minimal() +
  labs(colour = "Subject")


ggplot(data, aes(x = Z, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_line() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y vs. Z by Subject with Lines") +
  theme_minimal() +
  labs(colour = "Subject")

```

```{r Residual Plot}

data$index <- 1:nrow(data)
ggplot(data, aes(x = index, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject") 


ggplot(data, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject") 

```



# GCM based on LME

```{r}
pvals_gcm_lme_1 <- gcm.test(resid.XonZ = residuals_X_lme_1, resid.YonZ = residuals_Y_lme_1)$p.value

# Interpretation
cat("Based on the performed test based on LME:\n")
cat("- With a p-value of", pvals_gcm_lme_1, "and a significance level of 0.05,\n")
if (pvals_gcm_lme_1 < 0.05) {
  cat("- We reject the null hypothesis.\n")
} else {
  cat("- We fail to reject the null hypothesis.\n")
}
```
# Nonlinear model with random intercept

X_it = f(Z_it)+alpha_i+e_it
Y_it = g(Z_it)+beta_i+e_it


```{r Nonlinear dataset}
library(MASS) # For mvrnorm
library(tidyverse)
set.seed(123)

b <- 0

n_subjects <- 100
n_measurements <- 3

cor_matrix <- matrix(0.5, nrow = n_measurements, ncol = n_measurements)
diag(cor_matrix) <- 1

data <- data.frame(subject = rep(1:n_subjects, each = n_measurements),
                   measurement = rep(1:n_measurements, times = n_subjects),
                   Z = NA)

# Generate Z values
for (i in 1:n_subjects) {
  data$Z[data$subject == i] <- mvrnorm(n = 1, mu = rep(0, n_measurements), Sigma = cor_matrix)
}

# Initialize X and Y with NA values
data$X <- NA
data$Y <- NA

# # Function to calculate X based on current and previous Z
# calculate_X_Y <- function(Z_current, Z_previous, is_X = TRUE) {
#   # Handle the case for the first measurement
#   Z_previous <- ifelse(is.na(Z_previous), Z_current, Z_previous)
#   
#   if(is_X) {
#     return(5 * sin((Z_current + Z_previous)^2 + 10))
#   } else {
#     return((Z_current + Z_previous)^2)
#   }
# }


alpha <- rnorm(n_subjects,sd=5)
alpha <- rep(alpha, each = n_measurements)

beta <- rnorm(n_subjects,sd=5)
beta <- rep(beta, each = n_measurements)

# alpha <- rep(1:10, n_subjects)
# alpha <- rep(alpha, each = n_measurements)  # Fixed intercept for X
# beta <- rep(1:10, n_subjects)
# beta <- rep(beta, each = n_measurements)  # Fixed intercept for X

# Iterate through data to calculate X and Y based on Z_t and Z_t-1
for(i in 1:nrow(data)) {
  Z_current <- data$Z[i]
  Z_previous <- ifelse(data$measurement[i] == 1, NA, data$Z[i-1])
  # data$X[i] <- calculate_X_Y(Z_current,0, TRUE) + alpha[i] + rnorm(1, mean = 0, sd = 1)
  # data$Y[i] <- calculate_X_Y(Z_current,0, FALSE) + beta[i] + b*data$X[i]  +rnorm(1, mean = 0, sd = 1)
  data$X[i] <- log(Z_current^2) + alpha[i] + rnorm(1, mean = 0, sd = 1)  # Quadratic effect of Z
  data$Y[i] <- Z_current^3 + 2*beta[i] + rnorm(1, mean = 0, sd = 1)  # Sinusoidal effect of Z
  # data$X[i] <- log(Z_current^2) + alpha[i] + rnorm(1, mean = 0, sd = 1)  # Quadratic effect of Z
  # data$Y[i] <- Z_current^3 + 2*beta[i] + rnorm(1, mean = 0, sd = 1)  # Cubic effect of Z
}
  
```

# Check data structure

```{r Plot of X,Y on Z}
library(ggplot2)


ggplot(data, aes(x = Z, y = X)) +
  geom_point(aes(color = Z), alpha = 0.6) +  
  scale_color_gradient(low = "blue", high = "red") + 
  theme_minimal() +  
  labs(title = "Relationship between Z and X",
       x = "Z",
       y = "X") +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Z, y = Y)) +
  geom_point(aes(color = Z), alpha = 0.6) +  
  scale_color_gradient(low = "green", high = "yellow") + 
  theme_minimal() +  
  labs(title = "Relationship between Z and Y",
       x = "Z",
       y = "Y") +
  theme(plot.title = element_text(hjust = 0.5))  
```

```{r Average X and Y vs. Average Z 2}
# Use tidyverse to calculate averages and plot
data %>%
  group_by(subject) %>%
  summarise(average_X = mean(X, na.rm = TRUE), average_Z = mean(Z, na.rm = TRUE), average_Y = mean(Y, na.rm = TRUE)) %>%
  ggplot() +
  geom_point(aes(x = average_Z, y = average_X), color = "blue") +
  geom_point(aes(x = average_Z, y = average_Y), color = "red") +
  labs(x = "Average Z", y = "Average", title = "Average X and Y vs. Average Z for each subject") +
  theme_minimal() +
  scale_y_continuous(name = "Average X", sec.axis = sec_axis(~., name = "Average Y"))
```
```{r Relationship between Z and X for Selected Subjects 2}

library(ggplot2)

set.seed(123)
selected_subjects <- sample(unique(data$subject), 10)
selected_data <- data[data$subject %in% selected_subjects,]


ggplot(selected_data, aes(x = Z, y = X, group = subject, color = as.factor(subject))) +
  geom_line() +  
  geom_point() +  
  theme_minimal() +
  labs(title = "Relationship between Z and X for Selected Subjects",
       x = "Z",
       y = "X",
       color = "Subject") +
  scale_color_viridis_d(name = "Subject ID") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 10))

ggplot(selected_data, aes(x = Z, y = Y, group = subject, color = as.factor(subject))) +
  geom_line() +  
  geom_point() +  
  theme_minimal() +
  labs(title = "Relationship between Z and Y for Selected Subjects",
       x = "Z",
       y = "Y",
       color = "Subject") +
  scale_color_viridis_d(name = "Subject ID") +
  theme(legend.title = element_text(size = 12), legend.text = element_text(size = 10))

```
# Generalized addtiver regression

```{r Generalized addtiver regression 2 }
gam(X ~ s(Z), data = data, method = "GCV.Cp")


gam(Y ~ s(Z), data = data, method = "GCV.Cp")


ggplot(data, aes(x = Z, y = X)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "gam", formula = y ~ s(x), method.args = list(family = "gaussian")) +
  ggtitle("Relationship between X and Z") +
  theme_minimal()

ggplot(data, aes(x = Z, y = Y)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "gam", formula = y ~ s(x), method.args = list(family = "gaussian")) +
  ggtitle("Relationship between Y and Z") +
  theme_minimal()
```
```{r Residuals2}
data <- data %>% arrange(subject, Z)
data$residuals_x <- residuals(gam_model_x)
data$residuals_y <- residuals(gam_model_y)

ggplot(data, aes(x = Z, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_line() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X vs. Z by Subject with Lines") +
  theme_minimal() +
  labs(colour = "Subject")


ggplot(data, aes(x = Z, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_line() + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y vs. Z by Subject with Lines") +
  theme_minimal() +
  labs(colour = "Subject")

```
```{r Residual plot2}
data$index <- 1:nrow(data)
ggplot(data, aes(x = index, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject") 


ggplot(data, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject") 

```


```{r}
# 假设 res_X_gamm 和 res_Y_gamm 已经被计算出来
# 并且 adnidd$RID 包含了对应的主题标识（在你的示例中称为 subject）

# 为 X 的残差创建数据框
data_X <- data.frame(
  index = 1:length(res_X_gamm[,1]), # 假设对于第一种测量的残差
  residuals_x = res_X_gamm[,1], # 只取第一列作为示例
  subject = adnidd$RID
)

# 为 Y 的残差创建数据框
data_Y <- data.frame(
  index = 1:length(res_Y_gamm),
  residuals_y = res_Y_gamm,
  subject = adnidd$RID
)


```

# plot1
```{r}
library(ggplot2)


ggplot(data_X, aes(x = index, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none") 


ggplot(data_Y, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none") 


```
# plot2
```{r}
data_X_gam <- data.frame(
  index = 1:length(res_X_gam[,1]), # only the first col for plotting
  residuals_x = res_X_gam[,1], # only the first col for plotting
  subject = adnidd$RID
)


data_Y_gam_nobin <- data.frame(
  index = 1:length(res_YallZ),
  residuals_y = res_YallZ,
  subject = adnidd$RID
)

data_Y_gam_bin <- data.frame(
  index = 1:length(res_YbinZ),
  residuals_y = res_YbinZ,
  subject = adnidd$RID
)

ggplot(data_X_gam, aes(x = index, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for X1 by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none") 


ggplot(data_Y_gam_nobin, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y by Sample Index from GAM") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none") 



ggplot(data_Y_gam_bin, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Residuals for Y by Sample Index from GAM with binomial family") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none") 



```




```{r}
set.seed(123)
subjects_sampled <- sample(unique(data_X$subject), 100) #

data_X_sampled <- data_X[data_X$subject %in% subjects_sampled, ]
data_Y_sampled <- data_Y[data_Y$subject %in% subjects_sampled, ]


ggplot(data_X_sampled, aes(x = index, y = residuals_x, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Sampled Residuals for X by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none")


ggplot(data_Y_sampled, aes(x = index, y = residuals_y, colour = factor(subject))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggtitle("Sampled Residuals for Y by Sample Index") +
  theme_minimal() +
  labs(colour = "Subject")+
  theme(legend.position = "none") 

```


```{r ACF1}
res_x <- comp.resids(data$X, data$Z, regr.pars = list(), regr.method = "gam")
res_y <- comp.resids(data$Y, data$Z, regr.pars = list(), regr.method = "gam")

plot(res_x)
plot(res_y)
abline(h = 0, col = "red")
acf(res_y)
abline(h = 0, col = "red")
acf(res_x)
cor(res_x,res_y)
```

# GCM based on GAM for nonlinear case
```{r GCM based on GAM}
pvals_gcm <- gcm.test(data$X, data$Y, data$Z, regr.method = "gam", plot.residuals = F, regr.pars = list())$p.value
pvals_gcm
```
# GCM based on GAMM

```{r GCM based On GAMM}
model_X_gam_1 <- gamm4(X ~ s(Z), data = data, random = ~(1|subject))
model_Y_gam_1 <- gamm4(Y ~ s(Z), data = data, random = ~(1|subject))
residuals_X_gam_1 <- resid(model_X_gam_1$mer)
residuals_Y_gam_1 <- resid(model_Y_gam_1$mer)
pvals_gcm_gam_1 <- gcm.test(resid.XonZ = residuals_X_gam_1, resid.YonZ = residuals_Y_gam_1)$p.value
pvals_gcm_gam_1
```
```{r ACF}
plot(residuals_X_gam_1)
abline(h = 0, col = "red")
acf(residuals_X_gam_1)
plot(residuals_Y_gam_1)
abline(h = 0, col = "red")
acf(residuals_Y_gam_1)
```
```{r GAMM Predictions for X}

library(gamm4)
library(ggplot2)
library(dplyr)


predict_data_X <- expand.grid(
  Z = seq(min(data$Z), max(data$Z), length.out = 100),
  subject = factor(unique(data$subject))  # 
)


predict_data_X$predicted_X <- predict(model_X_gam_1$gam, newdata = predict_data_X)


ggplot(data, aes(x = Z, y = X, colour = factor(subject))) +
  geom_point(alpha = 0.5) + 
  geom_line(data = predict_data_X, aes(x = Z, y = predicted_X, group = subject), colour = "black", size = 1) +
  labs(title = "GAMM Predictions for X", x = "Z", y = "X") +
  theme_minimal()
```
```






